### 2011-06-10 standing-patches-paper.tex
/* Paper */ added some migration text

@@

patches where selection is $1+s_b$ separated by $1-s_d$ -- separation large relative to migration.   
(or relative to $s_d/m$?)

@@

Get migration and mutation rates from a homogeneous continuous model:
The migration rate from patch $i$ to a patch at distance $r$ is $C_i \exp(-\gamma r \sqrt{s} /\sigma)$,
where $C_i$ depends on the size of the patch and (to a lesser degree?) the local structural details of the selective difference.
The mutation rate is more or less (size of patch) times $\mu$; 
more accurately we can integrate using Barton assuming step selection.

>>>>

\section{Local establishment}

Include stuff from ``comparison to panmixia'' here.

@@

In this section, we suppose that instead of a uniform habitat,
that the selective pressure is {\em patchy}, so that the new alleles are only advantageous in isolated locations, 
and are disadvantageous in the intervening region.
As before, we assume that the new allele has fitness $1+s_b$ relative to the unmutated type in the ``good'' patches,
and assume that it has fitness $1+s_m$ in the intervening areas, with $s_m<0$.
To make things tractable, we assume that the distances between the patches are large relative to migration distance $\sigma$,
and that the disadvantage $s_m$ ensures that on the relevant timescale,
the allele is very unlikely to fix in the regions where it is at a disadvantage.
With these assumptions as justification, we will move to a model of discrete demes exchanging migrants,
where ``migration'' occurs through the rare, transient passage of the new allele through regions where it is disadvantageous.

equal deme sizes?  general migration rates?

@@

\subsection{Mutational influx}

Consider first a single, isolated patch of area $A$ in which no new allele has appeared.
\citet{barton1987} (or \citet{slatkin1981}) has studied this situation (in one dimension)
and found that:
mutations appearing further than a few multiples of $\sigma/\sqrt{2s_m}$ away from the patch have vanishingly small probability of establishment;
mutations appearing within the patch have probability strictly less but approaching that $2s_b$ of establishment;
and if the patch width is less than $2 \tan^{-1} (\sqrt{s_m})$, then local fixation does not occur (rather, is not stable).
Let $p(x)$ be the probability that a new allele arising at location $x$ fixes.
The function $p(x)$ can be found in terms of Jacobi elliptic functions,
XXX as shown in the Appendix,
but the expressions are particularly unhelpful.
The total successful mutational influx per generation is then given by $\int \rho \mu p(x) dx$,
and depends in a complicated way on the relationship of patch width and selection coefficients,
but still scales linearly with the population mutational influx $\rho \mu$.
Therefore, we define $p_f = \int p(x) dx$, and in a pinch will approximate this by $2 s_b A$,
which will be a good approximation if the width $\sqrt{A}$ is large relative to $\sigma/\sqrt{2s_m}$.

XXX We are interested in the case where a single mutation only takes over? XXX

\subsection{Migration rates}

Now suppose that there are two patches of area $A$ of good environment separated by distance $R$,
and that the allele has begun to fix in the first, but has not yet appeared in the second.
As above, any copy of the new allele that appears at location $x$ sufficiently near to the second patch 
has probability approximately $p(x)$ of fixing locally in this patch,
whether the allele is a new mutation or is descended from the occupants of the first patch.
There will be a small number of alleles descended from those of the first patch present in the regions where they are disadvantageous,
held at mutation--selection balance;
the equilibrium frequency may again be expressed in terms of Jacobi elliptic functions (Appendix XXX),
although the equations are again not particularly helpful.
However, we do know that if $q(r)$ is the expected proportion present at mutation--selection balance at distance $r$ from the first patch
in the absence of the second patch, that for large $r$ the proportion decays exponentially, $q(r) \approx C \exp( -|s_m r / \sigma|)$,
where $C$ is a constant depending on the structure of the populations and the selection coefficients.

Combining this with the calculations above allows us to get a rough sense of the scale of migration between separated patches,
supposing that patches are large enough for local fixation to be stable.
If the alleles were not advantageous in the second patch,
the average numbers of mutant alleles present near the second patch would be around $\rho A \exp( - |s_m R/\sigma| )$;
so the waiting time until some migrant allele fixes in the second patch would be something like $T_m = ( 2 s_b \rho A \exp( - |s_m R/\sigma|)^{-1}$.
If $T_m$ is large, then we would expect successful migrants to be rare,
so that if the second patch adapts through migration, it is likely to come from only a single migrant.
In the converse situation, XXX more than one migrant? XXX

Get migration and mutation rates from a homogeneous continuous model:
The migration rate from patch $i$ to a patch at distance $r$ is $C_i \exp(-\gamma r s /\sigma)$,
where $C_i$ depends on the size of the patch and (to a lesser degree?) the local structural details of the selective difference.
The mutation rate is more or less (size of patch) times $\mu$; 
more accurately we can integrate using Barton assuming step selection.
